---
title: "Week3_Assignment"
author: "B. Sosnovski"
date: "9/15/2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)

```

# Regular expressions and string processing

In this assignment, solutions to problems 3, 4 and 9 in Chapter 8 from textbook "Automated Data Collection with R" by S.Munzert, C. Rubba, P. Meißner, and D. Nyhuis, are provided below.

## Load packages
```{r}
library(stringr)
```

## Problem 3
```{r }
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5543642Dr. Julius Hibbert"

names <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
names           
```

### Part 1: Rearrange name vector into first_name last_name format
The following could be done without using functions because the vector containg the names is very short. However, if the vector had a relatively large length with many changes to be done, doing the same task without functions would be very repetitive.

As practice of using functions in R, I will use the function grepall as described in section 8.2 from the textbook and create another function to change the displayed order of a full name.

```{r}
grepall <- function(pattern, x, 
                    ignore.case = FALSE, perl = FALSE, 
                    fixed = FALSE, useBytes = FALSE, 
                    value = FALSE, logic = FALSE){
        # error and exception handling
        if(length(pattern) == 0 | length(x) == 0){
                warning("Length of pattern or data equals zero.")
                return(NULL)
        }
        # apply grepl() and all()
        indices <- sapply(pattern, grepl, x, ignore.case, perl, fixed, useBytes)
        index <- apply(indices, 1, all)
        # indexation and return of results
        if(logic == T) return(index)
        if(value == F) return((1:length(x))[index])
        if(value == T) return(x[index])
}
```

Identify the names listed as last name, first name and change the order.
```{r}
pattern <-"[[:alpha:]]+, [[:graph:] ]+"
old <- grepall(pattern,names,value=T)
indices <- grepall(pattern,names,value=F)
old
indices
# function that change the order of names to first and last names
new <- function(input){
        n<-length(input)
        flname <- vector('character')
        for (i in 1:n){
                split_name <- unlist(str_split(input[i],", "))
                temp <- paste(split_name[2], split_name[1])
                flname <- c(flname, temp)
        }
        return(flname)
}

names[indices] <- new(old)
names

```

### Part 2: Logical vector indicating which of the characters have titles

```{r}
pattern <- "[ [:upper:]][[:lower:]]+\\."
title <-str_detect(names, pattern)
title
```
### Part 3: Logical vector indicating if any character has a middle name

```{r}
pattern <- "[:upper:]\\. [[:alpha:]]+ "
middle_name <-str_detect(names, pattern)
middle_name
```

## Problem 4

Description of the following regular expressions with examples:

1) [0-9]+\\$

The regular expression returns a substring consisting of at least one of digits from 0 to 9 followed by character $.

```{r}
text<- " This is an example where digits appear: 111, 045$, 6.99, $401, 76$, 432$00."
unlist(str_extract_all(text, '[[0-9]]+\\$'))
```

2) \\b[a-z]{1,4}\\b

\\b indicates word edge, and so the expression returns whole word(s)  consisting of 1 to 4 letters from a to z.

```{r}
text<- " This is another example t2o what such expression wil1 return."
unlist(str_extract_all(text, '\\b[a-z]{1,4}\\b'))
```

3) .*?\\.txt$

$ indicates end of a string, and the expression returns the shortest possible sequence of any character zero or more times followed by the sequence of characters ".txt" in the end of the a string.

```{r}
text<- c("mydata.txt","main_data.txt.","xxx-data.txt", "data.txt contains all data.")
unlist(str_extract_all(text, ".*?\\.txt$"))
```

4) .*?\\.txt$

$ indicates end of a string, and the expression returns the shortest possible sequence of any character zero or more times followed by sequence of characters ".txt" in the end of the a string.

```{r}
text<- c("mydata.txt","main_data.txt.","xxx-data.txt", "data.txt contains all data.")
unlist(str_extract_all(text, ".*?\\.txt$"))
```

5) \\d{2}/\\d{2}/\\d{4}

The expression returns a sequence of characters in the date format xx/xx/xxxx (e.g., 09/15/2018).

```{r}
text<- c("Dec 14, 2001","This assignment is due on 09/16/2018.","Jane's birthday is 10/10/2010.", "When is yor birthday?")
unlist(str_extract_all(text, "\\d{2}/\\d{2}/\\d{4}"))
```

6) <(.+?)>.+?</\\1>

The regular expression returns a sequence of characters that describes a html tag. For example, "\<email\>some_email_address_here\</email\>". In the regular expression "\\1" is backreferencing to the substring from the expression "(.+?)".

```{r}
text<- ("<html>
<body>
<h1>This is heading 1</h1>
<h2>This is heading 2</h2>
<h3>This is heading 3</h3>
<h4>This is heading 4</h4>
<h5>This is heading 5</h5>
<h6>This is heading 6</h6>
</body>
</html>")
unlist(str_extract_all(text, "<(.+?)>.+?</\\1>"))
```


## Problem 9

The secret message below is sequence of characters from th alpha (upper and lower), digit and punct classes. Let's first extract the elements of upper and lower alpha classes, separetely.

```{r}
msg <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"
```

```{r}
lower <- str_extract_all(msg,"[[a-z]+]")
lower
```

```{r}
upper <- str_extract_all(msg,"[[A-Z]]+")
upper
```

Bingo!! Now it is just to make it presentable. Note that in the original string each word is separated by a punctuation mark.

```{r}
meaning <- unlist(str_extract_all(msg,"[[A-Z]]+|[[:punct:]]"))
meaning

meaning <- str_c(meaning, collapse ="")
meaning <- str_replace_all(meaning, pattern = "\\.+", replacement =" ")
meaning
```