---
title: "Project 2 - Dataset 1"
author: "E. Azrilyan and B. Sosnovski"
date: "9/29/2018"
output: 
  html_document: 
    theme: cerulean
---

```{r setup, include=FALSE, warning = FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Dataset

The dataset contains information from the U.S. Department of Agriculture about the longitude and latitude, state, address, name, zip code, etc, of Farmers Markets in the United States. The data is considered to be wide format because, for example, the variable produce/product is structured as if every single type of produce/product is a variable.

# Tidying, transforming and analyzing Data

Information is read from a csv file into R. The data has a "wide" structure. The data is then made tidy and tranformed, and analysis is performed.

The goal is to use the tidyverse packages, we will provide alternative ways of perfoming the analyses in this project.

* Which of the states have most of farmers markets with organic produce and meat? 

* What categories of foods are predominantly present in farmers markets in the U.S and by States?

_____________________________________________________________________

## Load packages

```{r}
library(knitr)
library(kableExtra)# manipulate table styles
suppressMessages(library(tidyverse))
```

## Read the data

```{r}
df <- read.csv("Export.csv")
str(df, give.attr = F) # with not to show attributes as substructures. If T then str is really long... 
```

## Select data
For our analysis, not all variables in the data frame are needed. So we select the ones that will be used in the analysis.

```{r}
wide_df <- df %>% select(FMID,State, Organic:WildHarvested)
```


## Tidying the data

This data is considered to be wide format because the variables from Organic to WildHarvested are structured as if there were single variables.

We can reshape the data into the long format by taking the Organic:WildHarvested transform then to a single column called "items".

```{r}
long_df <- wide_df %>% gather(Items, Sell_Item,Organic:WildHarvested)
long_df[long_df == "-"] <- NA
kable(head(long_df)) %>% kable_styling(bootstrap_options = "striped", full_width = F)
```

## Analysis

* Analysis 1: Which of the states have most of farmers markets offering both organic and meat products? 

* Analysis 2: What categories of foods are predominantly present in farmers markets in the U.S and by States?

### Analysis 1

Using the original wide format data:

```{r}
data1 <- wide_df %>% group_by(State) %>% 
        select(State, Organic, Meat)  %>% 
        filter(Organic=="Y", Meat=="Y") %>% 
        count(State) %>% 
        #ungroup() %>% #if this line is commented,the class function will show "grouped_df", 
                      #but there is no difference in the results shown in the print displayed.
        arrange(desc(n))
class(data1)
print.data.frame(data1) # The function "print.data.frame" from base R, to show all rows.
```

Using the long format data:

```{r}
data2 <- long_df %>% 
        select(FMID,State, Items,Sell_Item)  %>% 
        filter(Items=="Organic"| Items=="Meat") 
kable(head(data2))%>% kable_styling(bootstrap_options = "striped", full_width = F)

data2 <-data2 %>% group_by(State) %>% 
        spread(Items,Sell_Item) %>%
        filter(Organic=="Y",Meat=="Y") %>% 
        count(State) %>% 
        arrange(desc(n))
print(data2, n=15) # The function "print" from base R, to show n rows.
```

##### Conclusion:

California and New York are the top states in raw number of farmers market that sells both organic and meat products. An interesting, analysis is to check the proportion to the state's population.



### Analysis 2

#### Using dplyr only:

We can filter the information needed and and count the number of times a particular produce or product appears in the data frame.

```{r}
df_items <- long_df %>% select(Items, Sell_Item)
freq_items <- df_items %>% group_by(Items) %>% 
        filter(Sell_Item=="Y") %>% 
        count(Sell_Item) %>% 
        arrange(desc(n))
kable(freq_items) %>% kable_styling(bootstrap_options = "striped", full_width = F)

```

#### Using table() and dplyr:
 
Alternately, we can create a frequency table to count the number of markets that sell a particular type of produce or product. 

This seems a less efficient approach because we have to perform more transformations to the data.

```{r}
 factor(df_items)
 tb <- table(df_items) 
 kable(head(tb))%>% kable_styling(bootstrap_options = "striped", full_width = F)
 tb <- as.data.frame(tb)
 str(tb)
 tb <- tb %>% filter(tb$Sell_Item=="Y") %>% select(Items, Freq) %>% 
         arrange(desc(Freq))
 kable(tb) %>% kable_styling(bootstrap_options = "striped", full_width = F)
```


#### Displaying the data

In addition to a frequency table, we can also graph a bar plot of the number of markets that sell a particular type of produce (or product).

```{r}
 barplot(tb$Freq, main = "U.S. Farmers Market",
 ylab = "Frequency",
 names.arg = tb$Items,
 las=2,
 col = "blue")
```


##### Conclusion:

Data shows that most of the markets sell vegetables and baked goods. 


