---
title: "Project 2 - Dataset 3"
author: "E. Azrilyan and B. Sosnovski"
date: "10/04/2018"
output: 
  html_document: 
    theme: cerulean
---

```{r setup, include=FALSE, warning = FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Dataset

The website has information about the movies spread in 8 webpages.

## Load packages
```{r}
library(knitr)
library(XML)
library(kableExtra)# manipulate table styles
suppressMessages(library(RCurl))
suppressMessages(library(tidyverse))
```


## Function that creates a vector with all links to be accessed to retrieve data

```{r}
start_url <- "https://www.boxofficemojo.com/alltime/world/?pagenum="
end_url <- "&p=.htm"
# function
pages <- function(n){
        urls <- vector('character')
        for (i in 1:n){
                temp <- str_c(start_url,i,end_url, collapse = "")
                urls <- c(urls, temp)
        }
        return(urls)
}

urls <-pages(8)
urls
```

## Parsing data from web site

We creating a dataframe with empty vectors to storage all the information that is going to be read from the website.

```{r}
movie_df <- data.frame(V1 = numeric(0), V2 = character(0), V3 = character(0), V4 = numeric(0), V5 = numeric(0), V6 = numeric(0), V7 = numeric(0), V8 = numeric(0), V9=numeric(0), stringsAsFactors = FALSE)

sapply(movie_df, class)
```

Now we retrieve the information from each page, parse it and append to the above dataframe.

Tyding the data will also be performed in this step, since the numerical data contains the symbols $, % and ^.

```{r}
for (i in 1:8){
        url <-urls[i]
        htmlData <- getURL(url)
        parsedData <-htmlParse(htmlData, encoding = "UTF-8")
        data <- readHTMLTable(parsedData, skip.rows =1 )
        table <- data[[2]]
        # Converting the type of columns 1 to character type:
        table[,1] <- suppressWarnings(as.numeric(table[,1])) 
        #Converting the type of columns 2 and 3 to character type:
        for(j in seq(2,3)){
                table[,j] <- as.character(table[,j])
        }
        # Removing the symbols $, % and ^ from the numerical columns and changing them to numeric type in R:
        for(j in seq(4,9)){
                table[,j] <- suppressWarnings(as.numeric(gsub("[\\$%\\^,]","", table[,j])))
        }
        movie_df <- rbind(movie_df, table, sort = F)

}
sapply(movie_df, class)
head(movie_df)
dim(movie_df)
```

Because some of the gross values are in the thousands (they appear on the webpages with k), R replace it with NAs.


Rename the columns of the movie_df:

```{r}
colnames(movie_df) <- c( "Rank", "Title" , "Studio", "Worldwide", "Domestic", "Domestic.Perc", "Overseas", "Overseas.Perc", "Year")

head(movie_df)
```


How to order the datafram?? 
```{r}
#new_df <- movie_df[order(movie_df$Rank),]


```
