---
title: "Week 7 Assignment"
author: "B. Sosnovski"
date: "10/13/2018"
output: 
  html_document: 
    theme: cerulean
---
# Assignment Overview

In this assignment, we are required to create three files in HTML table, XML, and JSON formats. The files contain information about three of our favorite books in a particular subject area, with at least one book having multiple authors. 

My choices don't necessarily are my favorite books since two of the chosen books were are excellent references in cryptography, my research field.

We use R to load the information from each of the three sources into separate R data frames. We investigate if the three data frames identical or not.

_____________________________________________________________________

```{r setup, include=FALSE, warning = FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages

```{r}
library(knitr)
library(kableExtra)# manipulate table styles
library(XML)
suppressMessages(library(RCurl))
library(jsonlite)
```

## Read and Parse Data


#### HTML

```{r}
html.url <- "https://raw.githubusercontent.com/bsosnovski/DATA607/master/Week7Assignment/mybooks.html"
htmlData <- getURL(html.url)
html_parsed <-htmlParse(htmlData, encoding = "UTF-8")
html_parsed
class(html_parsed)
htmlData <- readHTMLTable(html_parsed, stringsAsFactors = FALSE)
html_df <- htmlData[[1]] #without this codeline there is a 'NULL' printed in the first line of htmlTable.
class(html_df)
kable(html_df) %>% kable_styling(bootstrap_options = "striped")
```

#### XML

```{r}
xml.url <- "https://raw.githubusercontent.com/bsosnovski/DATA607/master/Week7Assignment/mybooks.xml"
xmlData <- getURL(xml.url)
xml_parsed <- xmlParse(xmlData)
xml_parsed
class(xml_parsed)

xml_df <- xmlToDataFrame(xml_parsed)
colnames(xml_df) <- c("ID", "Title", "Author 1", "Author 2","Publication Year", "Publisher", "Number of Pages")
class(xml_df)
kable(xml_df) %>% kable_styling(bootstrap_options = "striped")

```

#### JSON

```{r}
json.url <- "https://raw.githubusercontent.com/bsosnovski/DATA607/master/Week7Assignment/mybooks.json"
jsonData <- getURL(json.url)
prettify(jsonData)

json_parsed <- fromJSON(jsonData)
json_parsed
class(json_parsed)

json_df <- data.frame(json_parsed)
colnames(json_df) <- c("ID", "Title", "Author 1", "Author 2","Publication Year", "Publisher", "Number of Pages")
class(json_df)
kable(json_df) %>% kable_styling(bootstrap_options = "striped")

```

The dataframes had different columns names and after changing the column names they look similar. However, as R objects they are not identical. 

```{r}
identical(html_df,xml_df)
identical(html_df,json_df)
identical(xml_df, json_df)
```
